### 2022-03-09 paper/localpsf.tex
paper edits

\subsection{Overview of the method}
The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
\begin{equation*}
\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
\end{equation*}
denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
\begin{equation}
\label{eq:scaled_impulse_response}
\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
\end{equation}
denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
\begin{equation*}
\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
\end{equation*}
denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively. In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).

@@

\section{Method}
\label{sec:method}
In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$. This section is organized as follows:
\begin{description}
	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
\end{description}
The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}. The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.

@@

\section{Fast ellipsoid intersection test}
\label{sec:fast_ellipsoid_intersection_test}
The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_p$ and $E_q$ be the ellipsoids defined as
\begin{align*}
	E_p :=& \{x : (x - \spatialmean_p)^T \spatialcov_p^{-1} (x - \spatialmean_p) \le \tau^2\} \\
	E_q :=& \{x : (x - \spatialmean_q)^T \spatialcov_q^{-1} (x - \spatialmean_q) \le \tau^2\}, \\
\end{align*}
where $\spatialmean_p, \spatialmean_q \in \mathbb{R}^\gdim$, and $\spatialcov_p, \spatialcov_q \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite. Let $K$ be the following one dimensional convex function:
\begin{equation*}
	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_p - \spatialmean_q)^T \left(\frac{1}{1-s}\spatialcov_p + \frac{1}{s}\spatialcov_q\right)^{-1}(\spatialmean_p - \spatialmean_q)	
\end{equation*}
In \cite{ELLIPSOIDINTERSECT} it is shown that $E_p \cap E_q = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$. We check whether $E_p$ and $E_q$ intersect by minimizing $K(s)$ on $(0,1)$. If $K(s^*) <0$ at the minimizer $s^*$, then $E_p \cap E_q = \{\}$. Otherwise $E_p \cap E_q \neq \{\}$.

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
\begin{equation*}
	\spatialcov_p P = \spatialcov_q P \Lambda,
\end{equation*}
where $P \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $P$ simultaneously diagonalizes $\spatialcov_p$ and $\spatialcov_q$, in the sense that $P^T\spatialcov_p P = \Lambda$, and $P^T\spatialcov_q P = I$, where $I$ is the $\gdim \times \gdim$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
	\label{eq:Ks_generalized}
	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \Phi^T\left(\spatialmean_p - \spatialmean_q\right)$. We compute the generalized eigenvalue decomposition of $\spatialcov_p$ and $\spatialcov_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used). The resulting algorithm for checking whether $E_p$ and $E_q$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	\SetKwProg{Fn}{Function}{}{}
	\Input{Ellipsoid $E_p$ with mean $\spatialmean_p$ and covariance $\spatialcov_p/\tau^2$\\
		Ellipsoid $E_q$ with mean $\spatialmean_q$ and covariance $\spatialcov_q/\tau^2$ 
	}

	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_p \cap E_q \neq \{\}$ and false otherwise}

	Solve generalized eigenvalue problem $\spatialcov_p P = \spatialcov_q P \Lambda$

	$v \gets P^T\left(\spatialmean_p - \spatialmean_q\right)$

>>>>

\subsection{Outline of the method}
\label{sec:intro_overview}

The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
\begin{equation*}
\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
\end{equation*}
denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
\begin{equation}
\label{eq:scaled_impulse_response}
\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
\end{equation}
denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
\begin{equation*}
\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
\end{equation*}
denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively. In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).

@@

\section{Method}
\label{sec:method}
In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$, as outlined in Section \ref{sec:intro_overview}.
%The supports of the impulse responses $\phi_x$ are estimated to be contained within ellipsoids (Section \ref{eq:mean_and_covariance_estimation}). The shapes of the ellipsoids depend on spatially varying volume, mean, and covariance functions, $V,\mu,\Sigma$, which we calculate by applying $\Aop$ to a small number of polynomial functions (Section \ref{sec:moments}). We use the ellipsoid support estimates in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}). Batches, $\eta_b$, of impulse responses, $\phi_{x_i}$, are computed by applying $\Aop$ to Dirac combs associated with the batches of sample points $S_b$. The impulse response batch $\eta_b$ contains the information about all impulse responses $\phi_{x_i}$ associated with points $x_i \in S_b$. Hence, once a small number of impulse response batches $\eta_b$ are computed, they can be used to perform pointwise evaluations $z \mapsto \phi_{x_i}(z)$ for a large number of impulse responses, $\phi_{x_i}$ (Section \ref{sec:get_impulse_response}).
%
%In Section \ref{eq:mean_and_covariance_estimation} we describe our ellipsoid estimate for the support of $\phi_x$. The ellipsoid shape depends on the functions $V,\mu,\Sigma$. We show how to compute these functions in Section \ref{sec:moments}. The ellipsoid support estimates are then used in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
This section is organized as follows:
\begin{description}
	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
\end{description}
The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}. The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.

@@

\section{Fast ellipsoid intersection test}
\label{sec:fast_ellipsoid_intersection_test}
The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
\begin{align*}
	E_1 :=& \{x : (x - \spatialmean_1)^T \spatialcov_1^{-1} (x - \spatialmean_1) \le \tau^2\} \\
	E_2 :=& \{x : (x - \spatialmean_2)^T \spatialcov_2^{-1} (x - \spatialmean_2) \le \tau^2\}, \\
\end{align*}
where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite. Let $K$ be the following one dimensional convex function:
\begin{equation*}
	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2)	
\end{equation*}
In \cite{ELLIPSOIDINTERSECT} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$. We check whether $E_1$ and $E_2$ intersect by minimizing $K(s)$ on $(0,1)$. If $K(s^*) <0$ at the minimizer $s^*$, then $E_1 \cap E_2 = \{\}$. Otherwise $E_1 \cap E_2 \neq \{\}$.

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
\begin{equation*}
	\spatialcov_1 P = \spatialcov_2 P \Lambda,
\end{equation*}
where $P \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $P$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $P^T\spatialcov_1 P = \Lambda$, and $P^T\spatialcov_2 P = I$, where $I$ is the $\gdim \times \gdim$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
	\label{eq:Ks_generalized}
	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \Phi^T\left(\spatialmean_1 - \spatialmean_2\right)$. We compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used). The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	\SetKwProg{Fn}{Function}{}{}
	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
	}

	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}

	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$

	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
